### 2021-09-14 Code/Manuscript-SMOTE.Rmd

---
title: "Host richness increases the occurrence but not the severity of bark beetle-induced tree mortality"
author: "Sarah J. Hart"
email: "sarah.hart@colostate.edu"
date: '`r format(Sys.time(), "%B %d, %Y")`'
output: 
  bookdown::word_document2:
    reference_docx: WordTemplate.docx
    fig_caption: yes
    toc: no
    number_sections: no
    df_print: kable
csl: "`r here:::here('ecology.csl')`"
bibliography: "`r here:::here('MyLib.bib')`"
urlcolor: blue
linkcolor: blue
citationcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = F,
	message = FALSE,
	warning = FALSE,
	dpi = 300,
	progress = FALSE,
	cache = FALSE
)

ipak <- function(pkg){
    new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
    if (length(new.pkg)>0) {
             install.packages(new.pkg, dependencies = TRUE)
    }
    easypackages::libraries(pkg)
}

options(repos = c(CRAN = "http://cran.rstudio.com"))

### Import libraries
ipak(c('tidyverse', 'tidymodels', 'reshape2', 'here', 'rgdal', 'sp', 'raster', 'stars', 'sf', 'prism', 'tigris', "tmap", 'fasterize', 'ggplot2','RColorBrewer','ggsci', 'patchwork', 'colorspace', 'grid', 'knitr',  'kableExtra', 'bookdown',  'pdp', 'ranger', 'VSURF', 'themis', 'foreach'))

#foreach', 'parallel', 'doParallel', 'quickPlot', 'reshape2','vegan', 'FSA', 
# Set seed
set.seed(2020)

# Set custom plotting theme
theme_new <- function(base_size = 9,base_family = "Helvetica"){
  theme_classic(base_size = base_size, base_family = base_family) %+replace%
    theme(
      axis.line.x = element_line(color="black", size = 0.25),
      axis.line.y = element_line(color="black", size = 0.25),
      axis.title = element_text(size = 9),
      axis.text = element_text(colour="black", size=8),
      legend.key=element_rect(colour=NA, fill =NA),
      panel.grid = element_blank(),   
      plot.background = element_rect(fill = NA, colour = NA),
      panel.border = element_rect(fill = NA, colour = NA),
      panel.background = element_rect(fill = "white", colour = "black"), 
      strip.background = element_rect(fill = "white"),
      strip.text = element_text(size = 9)

    )
}
theme_black = function(base_size = 12, base_family = "Helvetica") {

  theme_grey(base_size = base_size, base_family = base_family) %+replace%

    theme(
      # Specify axis options
      axis.line = element_blank(),  
      axis.text.x = element_text(size = base_size*0.8, color = "white", lineheight = 0.9),  
      axis.text.y = element_text(size = base_size*0.8, color = "white", lineheight = 0.9),  
      axis.ticks = element_line(color = "white", size  =  0.2),  
      axis.title.x = element_text(size = base_size, color = "white", margin = margin(10, 0, 0, 0)),  
      axis.title.y = element_text(size = base_size, color = "white", angle = 90, margin = margin(0, 10, 0, 0)),  
      axis.ticks.length = unit(0.3, "lines"),   
      # Specify legend options
      legend.background = element_rect(color = NA, fill = rgb(50, 50, 50, maxColorValue = 255)),  
      legend.key = element_rect(color = "white",  fill = rgb(50, 50, 50, maxColorValue = 255)),  
      legend.key.size = unit(1.2, "lines"),  
      legend.key.height = NULL,  
      legend.key.width = NULL,      
      legend.text = element_text(size = base_size*0.8, color = "white"),  
      legend.title = element_text(size = base_size*0.8, face = "bold", hjust = 0, color = "white"),  
      legend.position = "right",  
      legend.text.align = NULL,  
      legend.title.align = NULL,  
      legend.box = NULL, 
      # Specify panel options
      panel.background = element_rect(fill = rgb(50, 50, 50, maxColorValue = 255), color  =  NA),  
      panel.border = element_rect(fill = NA, color = "white"),  
      panel.grid.major = element_line(color = NA),  
      panel.grid.minor = element_line(color = NA),  
      panel.margin = unit(0.5, "lines"),   
      # Specify facetting options
      strip.background = element_rect(fill = "grey30", color = "grey10"),  
      strip.text.x = element_text(size = base_size*0.8, color = "white"),  
      strip.text.y = element_text(size = base_size*0.8, color = "white",angle = -90),  
      # Specify plot options
      plot.background = element_rect(color =rgb(50, 50, 50, maxColorValue = 255), fill = rgb(50, 50, 50, maxColorValue = 255)),  
      plot.title = element_text(size = base_size*1.2, color = "white"),  
      plot.margin = unit(rep(1, 4), "lines")

    )

}
theme_set(theme_new())

options(ggplot2.continuous.colour="viridis")
scale_colour_discrete <- scale_fill_npg
scale_fill_discrete <- scale_fill_npg
options(scipen=999)

p1col=c(rgb(104, 17, 10,  maxColorValue = 255), rgb(79, 130, 156, maxColorValue = 255), rgb(243, 173, 1, maxColorValue = 255), rgb(112, 146, 115,  maxColorValue = 255), rgb(112, 146, 115, maxColorValue = 255))

p1fill=c(rgb(104, 17, 10, alpha=230, maxColorValue = 255), rgb(79, 130, 156, alpha=230, maxColorValue = 255), rgb(243, 173, 1, alpha=50, maxColorValue = 255), rgb(112, 146, 115, alpha=230,  maxColorValue = 255), rgb(112, 146, 115, alpha=230, maxColorValue = 255))

p2col =c(rgb(20,20,20, maxColorValue = 255), rgb(54, 124, 111, maxColorValue = 255), rgb(235, 137, 52, maxColorValue = 255), rgb(83, 74, 90, maxColorValue = 255), rgb(232, 177, 60, maxColorValue = 255),rgb(90, 129, 154, maxColorValue = 255),rgb(113, 0, 2, maxColorValue = 255), rgb(166, 166, 166,  maxColorValue = 255))

fig.width1 <- 3.4
fig.width2 <- 7
fig.width1.5 <- 4.48

pt.width1 <-fig.width1 * 72
pt.width2 <-fig.width2 * 72
pt.width1.5 <- fig.width1.5 * 72

# Set projection - NAD83 / Conus Albers
aea.proj <- "EPSG:5070"

options(timeout=60*30) #timeout downloads that last longer than 30 minutes

# Set directory structure for project
dir.create(here("Data/"), showWarnings = FALSE)
dir.create(here("Results/"), showWarnings = FALSE)
dir.create(here("Results/Figures/"), showWarnings = FALSE)
```

# Abstract

# Keywords

# Introduction

Paragraph 1: Broad applied context

In this context there is a particular need to understand interactions between outbreaks of irruptive insects, which in recent years (2003-2012) have affected 85 million hectares of forest globally, or ca. 18 million hectares more than wildfire (van Lierop et al. 2015).

- insect outbreaks are causing widespread tree mortality in temperate forest worldwide
- future changes in climate are expected to increase tree mortality
- forests provide key ecosystem services
- management

Paragraph 2: Theoretical context - diversity & natural enemies
- Many studies have sought to understand how community diversity influences interactions between natural enemies and their resources.
- Key hypotheses:
1 - associational resistance - the abundance or damage of herbivores is lower when plant diversity is greater [@barbosa_associational_2009, @jactel_tree_2021]
hypotheses:
1 - resource concentration: the presence of heterospecific neighbors around a tree of a focal species grown in mixed stands leads to a lower probability of host tree finding by insect herbivores due to lower host abundance or frequency [@root_organization_1973]
2 - host apparency: 
3 - decoy: preference for non-host trees

However, benefits provided by mixtures are less evident for larger-scale disturbances [@jactel_tree_2021]

suggesting that changes in the structure of host communities, rather than biodiversity per se, can explain when a dilution effect should be observed.

Most studies have focused on individual- or population-level outcomes.

For instance, increased host community diversity may either increase or decrease the susceptibility of individuals to their natural enemies.

At the population-scale these effects may either scale-up or result in counter-intuitive effects. However, little research has examined community-level outcomes, which are often hard to quantify because of the number of potential interactions increases dramatically with increasing community diversity, particularly when natural enemies are generalists. Further, the effects of different natural enemies on their focal host populations often differ greatly, due to differences in susceptibility and mortality rates and the availability and quality (as viewed by their natural enemies) of resource communities. Critically, community-level outcomes may drive ecosystem processes, particularly when community diversity is low and natural enemy-resource relationships are highly dynamic.

*** paragraph about irruptive species and community host diversity*** 
- stand scale
- landscape scale

To better understand how community diversity influences interactions between natural enemies and their resources, we use a natural system with inherently low resource and natural enemy diversity. Specifically our research focuses on tree mortality due to three bark beetle species, the mountain pine beetle (*Dendroctonus ponderosae*), spruce beetle (*D. rufipennis*), and western balsam bark beetle (*Dryocoetes confusus*), which in subalpine forests of Western North America predominantly attack lodgepole pine (*Pinus contorta*), Engelmann spruce (*Picea engelmannii*), and subalpine fir (*Abies lasiocarpa*), respectively. We use this system to ask whether host tree richness or identity influence the occurrence and severity of tree mortality due to bark beetles at the community-scale? We hypothesize that
greater host richness will increase the probability of outbreak occurrence

Bark beetles (Curculionidae: Scolytinae) are among the few native insect species that can kill large numbers of trees in a single year. Bark beetles bore through the bark, where they mate and oviposit their eggs. Concurrently, bark beetles introduce pathogenic fungi Larvae feeding upon the phloem and fungal spread stop the translocation of water and nutrients and cause tree death. Conifer defense against bark beetles consists primarily of resin exudation that physically expels the beetle and allelochemicals, which repel or kill beetles. To overcome these defenses and colonize live trees, bark beetles rely on a mass-attack strategy, where pioneering beetles emit aggregation pheromones that call conspecifics to the tree. Typically bark beetles exist at low population levels and attack weakened trees, but as populations increase bark beetles attack increasingly better defended trees.For instance, in the continental western United States bark beetles have killed more than XXXX trees over the past X decades. Such severe mortality occurs only during outbreaks when pheromone-mediated mass-attack allows bark beetles to overcome tree defenses. 
For instance, a primary goal of the Western Bark Beetle is to promote resilience of forests to bark beetle outbreaks by increasing the diversity of age classes and tree species.

If greater host richness increases niche space for bark beetles, then community-level occurrence of bark beetle activity will be greater in more diverse communities. community-level severity, here the amount and proportion of basal area affected by bark beetles, will be lower in stands with more host species.

To test these hypotheses, we used a large dataset consisting of XXX,XXX plots established by the United States Forest Service Forest Inventory and Analysis Program (FIA; https://www.fia.fs.fed.us/). Specifically we ask: (1) Is the occurrence or severity of bark beetle

# Methods

## Study area
```{r set species and agent codes}
# SPECIES CODES

# MPBhosts
PICO <- 108
PIAL <- 101
PIAR <- 102
PIFL <- 113
PIST <- 114
PIMO <- 119
PIPO <- 122

# SBhosts
PIEN <- 93
PIPU <- 96

#SFDhosts
ABLA <-19 
ABCO <- 15
ABGR <- 17

ohostsMPB <- c( PIAL, PIAR, PIFL, PIST, PIMO, PIPO)
ohostsWBBB <- c(ABCO, ABGR)
ohostsSB <- PIPU
otherhosts <- c(ABCO, ABGR, PIPU, PIAL, PIAR, PIFL, PIST, PIMO, PIPO)

# AGENT CODES
# DCA
bb <- 11000 #bark beetle
mpb <- 11006 #mountain pine beetle
sb <- 11009 #spruce beetle
wbbb <- 11015 # western balsam bark beetle
sfm <- 80002 # subalpine fir mortality
```

The study area consists of subalpine lodgepole pine (*Pinus contorta*), subalpine fir (*Abies lasiocarpa*), and Engelmann spruce (*Picea engelmannii*) forest in the Intermountain West (i.e., Arizona, New Mexico, Colorado, Utah, Nevada, Idaho, Montana, Wyoming; Fig. \@ref(fig:FigSA)

## Data
### FIA data

The FIA program is a single inventory program that includes all public and private forested land (>= 0.4 ha in size and >= 10% canopy cover) in the US. In the Western US, all plots are visited once every ten years [@gray_forest_2012]. The spatially and temporally distributed probabilistic sampling design is useful studies of the distribution of tree species [e.g., @iverson_predicting_1998; @rehfeldt_empirical_2006], forest insects [e.g., @derose_effects_2013], and tree mortality [e.g., @shaw_forest_2005]

At each FIA plot, field crews collect data for trees (>= 12.7 cm DBH) within four 7.32 m radius subplots arranged in a fixed pattern. Data on the proximate cause of death is collected for any tree (>= 12.7 cm DBH) that was alive at the previous visit and at revisit is dead using visible evidence (e.g., fire scars, bark beetle galleries).

```{r download FIA data, eval=F}
dir.create(here("Data/"), showWarnings = FALSE)
dir.create(here("Data/FIA"), showWarnings = FALSE)
# reference data
dir.create(here("Data/FIA/Reference"), showWarnings = FALSE)
temp <- here("/Data/FIA/Reference/FIADB_REFERENCE.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/FIADB_REFERENCE.zip", temp)
unzip(temp, exdir=here("/Data/FIA/Reference"))

# AZ
dir.create(here("Data/FIA/AZ"), showWarnings = FALSE)
temp <- here("/Data/FIA/AZ/AZ.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/AZ.zip", temp)
unzip(temp, exdir=here("/Data/FIA/AZ"))

# CO
dir.create(here("Data/FIA/CO"), showWarnings = FALSE)
temp <- here("/Data/FIA/CO/CO.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/CO.zip", temp)
unzip(temp, exdir=here("/Data/FIA/CO"))

# ID
dir.create(here("Data/FIA/ID"), showWarnings = FALSE)
temp <- here("/Data/FIA/ID/ID.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/ID.zip", temp)
unzip(temp, exdir=here("/Data/FIA/ID"))

# MT
dir.create(here("Data/FIA/MT"), showWarnings = FALSE)
temp <- here("/Data/FIA/MT/MT.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/MT.zip", temp)
unzip(temp, exdir=here("/Data/FIA/MT"))

# NM
dir.create(here("Data/FIA/NM"), showWarnings = FALSE)
temp <- here("/Data/FIA/NM/NM.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/NM.zip", temp)
unzip(temp, exdir=here("/Data/FIA/NM"))

# NV
dir.create(here("Data/FIA/NV"), showWarnings = FALSE)
temp <- here("/Data/FIA/NV/NV.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/NV.zip", temp)
unzip(temp, exdir=here("/Data/FIA/NV"))

# UT
dir.create(here("Data/FIA/UT"), showWarnings = FALSE)
temp <- here("/Data/FIA/UT/UT.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/UT.zip", temp)
unzip(temp, exdir=here("/Data/FIA/UT"))

# WY
dir.create(here("Data/FIA/WY"), showWarnings = FALSE)
temp <- here("/Data/FIA/WY/WY.zip")
download.file("https://apps.fs.usda.gov/fia/datamart/CSV/WY.zip", temp)
unzip(temp, exdir=here("/Data/FIA/WY"))
```

```{r download spatial data, eval=F}
# states
dir.create(here("Data/Spatial/"), showWarnings = FALSE)
dir.create(here("Data/Spatial/States"), showWarnings = FALSE)
temp <- here("/Data/Spatial/States/cb_2018_us_state_20m.zip")
download.file("https://www2.census.gov/geo/tiger/GENZ2018/shp/cb_2018_us_state_20m.zip", temp)
unzip(temp, exdir=here("/Data/Spatial/States"))

# usfs western bark beetle strategy
dir.create(here("Data/Spatial/USFS/"), showWarnings = FALSE)
temp <- here("/Data/Spatial/USFS/WBBS_PL.gdb.zip")
download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/fc/S_USA.Activity_WBBS_PL.gdb.zip", temp)
unzip(temp, exdir=here("/Data/Spatial/USFS/"))

# usfs timber harvest
dir.create(here("Data/Spatial/USFS/"), showWarnings = FALSE)
temp <- here("/Data/Spatial/USFS/Activity_TimberHarvest.gdb.zip")
download.file("https://data.fs.usda.gov/geodata/edw/edw_resources/fc/S_USA.Activity_TimberHarvest.gdb.zip", temp)
unzip(temp, exdir=here("/Data/Spatial/USFS/"))
```

```{r import raw data, eval=F}
wd <- here("Data/FIA")

# List directories containing data for each state
dirs <- list.dirs(wd)

dirs <- dirs[c(-1, -8)]

# Create empty lists to hold dataframe for each state
dats <- list() 
datp <- list()
datc <- list()
datt <- list()
datspc <- list()

for(i in 1:length(dirs)){
  state <- strsplit(dirs[i], split="FIA/")[[1]][2]
  # Survey 
  dats[[state]] <- read.csv(paste0(dirs[i], "/",state, "_SURVEY.csv"), stringsAsFactors = F)
  # Plot
  datp[[state]] <- read.csv(paste0(dirs[i], "/",state, "_PLOT.csv"), stringsAsFactors = F)
  # Condition
  datc[[state]] <- read.csv(paste0(dirs[i], "/",state, "_COND.csv"), stringsAsFactors = F)
  # Tree
  datt[[state]] <- read.csv(paste0(dirs[i], "/",state, "_TREE.csv"), stringsAsFactors = F)
  # Subplot Condition
  datspc[[state]] <- read.csv(paste0(dirs[i], "/",state, "_SUBP_COND.csv"), stringsAsFactors = F)
}

### Merge data
datsm <- do.call(rbind, dats)
datpm <- do.call(rbind, datp)
datcm <- do.call(rbind, datc)
dattm<- do.call(rbind, datt)
datspcm<- do.call(rbind, datspc)

remove(datt, datp, datc, dats,datspc)

# Merge Subplot & Plot data
df <- merge(datcm[,c("PLT_CN", "CONDID","COND_STATUS_CD","OWNCD", "STDAGE")], datspcm[,c("PLT_CN", "CONDID", "SUBP")], by=c("PLT_CN", "CONDID"))

# Merge Survey & Plot data
datz <- merge(datsm[,c("CN", "ANN_INVENTORY", "STATEAB")], datpm[,c("CN","SRV_CN","INVYR", "COUNTYCD","PLOT", "PLOT_STATUS_CD", "KINDCD", "DESIGNCD", "MANUAL", "SAMP_METHOD_CD", "REMPER", "MEASYEAR", "ELEV", "LAT", "LON", "PREV_PLT_CN", "UNITCD")],  by.y="SRV_CN", by.x="CN", all==TRUE)
colnames(datz)[which(colnames(datz) %in% c("CN", "CN.y"))] <- c("SRV_CN", "PLT_CN")

#confirm each plot record is in data frame only once 
if(max(sort(table(datz$PLT_CN)))!=1){
  print("stop there are duplicated plot records")
}
datz.sp<-datz
```

```{r merge survey, plot, condition & tree data, eval=F}
## select annual inventory
datz <-datz[datz$ANN_INVENTORY=='Y',]; nrow.annual <- nrow(datz)

## select plots where all of the subplots are forested
allspforested <- unique(df$PLT_CN[!(df$PLT_CN %in% subset(aggregate(COND_STATUS_CD~PLT_CN, FUN=function(x){length(x[x!=1])}, df), COND_STATUS_CD>0)$PLT_CN)])
allspforested <- allspforested[allspforested %in% datz$PLT_CN]
datz <- datz[datz$PLT_CN %in% allspforested, ]

#Merge Survey-Plot-Condition & Tree data
dattm.sub <-subset(dattm, PLT_CN %in% datz$PLT_CN)
datz <- merge(datz, dattm.sub[, c("PLT_CN", "CN", "STATECD", "COUNTYCD","TREE", "DIST", "STATUSCD", "DIA", "SPCD","CCLCD", "AGENTCD", "DAMAGE_AGENT_CD1","DAMAGE_AGENT_CD2", "DAMAGE_AGENT_CD3", "TPA_UNADJ")],  by="PLT_CN", all=FALSE)
datz$PLOT.STATEAB <- paste(datz$PLOT, datz$STATEAB, sep=".")

# subset data to only trees greater than 5 inches (12.7 cm) DBH
datz5 <- subset(datz, DIA>=5)

# subset data to only trees within subplot
datz5 <- subset(datz5, DIST<=24)

# remove trees with missing values for TPA count 
datz5 <- datz5[is.na(datz5$TPA_UNADJ)==FALSE, ] ## not sure why missing values occur here
datz5.bak <- datz5

spruce.plots <-unique(datz5[datz5$SPCD == PIEN,"PLT_CN"])
pine.plots <-unique(datz5[datz5$SPCD == PICO,"PLT_CN"])
fir.plots <- unique(datz5[datz5$SPCD == ABLA,"PLT_CN"])
datz5 <- datz5[datz5$PLT_CN %in% c(spruce.plots, pine.plots, fir.plots),]

## subset to recently killed trees
datz5.bak <- datz5
datz5$STATUSCD <- factor(datz5$STATUSCD , levels=c(0:3), labels=c("No Status", "Live", "Dead", "Removed"))
datz5$STATUSCD.2<- ifelse(datz5$STATUSCD=="Dead" & is.na(datz5$AGENTCD) == FALSE, "Recently killed",as.character(datz5$STATUSCD))
datz5 <- datz5[datz5$STATUSCD.2 %in% c("Live", "Recently killed"),]
```

#### Determination of presence/absence of bark beetle activity

For each live tree (>= 12.7 cm DBH), field crews record up to three damaging agents, which are defined as agents that are likely to prevent the tree from surviving >2 years, reduce the growth of the tree in the near term, or negatively affect the tree's marketable products [@burrill_forest_2017].

Cause of death codes are very broad (e.g., "insect" or "disease"). Accuracy of FIA data is commonly assessed using blind checks, where two crews perform independent inventories. Agreement between mortality agent codes recorded in the two inventories is generally >80% [@anderegg_tree_2015]. Active damage is easier to identify, thus the codes for damaging agents are much more specific.

```{r determine presence of BB agents in focal hosts, eval=F}
datz5$WBBB.l <- NA
datz5$WBBB.d <- NA
datz5$WBBB.l <- ifelse(datz5$DAMAGE_AGENT_CD1==bb & datz5$SPCD==ABLA |datz5$DAMAGE_AGENT_CD1 %in% c(wbbb, sfm) & datz5$SPCD == ABLA, 1, 0) 
datz5$WBBB.d <- ifelse(datz5$AGENTCD==10 & datz5$SPCD == ABLA, 1, 0) 
datz5$WBBB <- rowSums(datz5[,c("WBBB.d", "WBBB.l")], na.rm=T)
datz5$WBBB[is.na(datz5$WBBB)==T]<-0

datz5$MPB.l <- NA
datz5$MPB.d <- NA
datz5$MPB.l <- ifelse(datz5$DAMAGE_AGENT_CD1==bb & datz5$SPCD ==PICO | datz5$DAMAGE_AGENT_CD1 ==11006 & datz5$SPCD ==PICO, 1, 0) 
datz5$MPB.d <- ifelse(datz5$AGENTCD==10 & datz5$SPCD ==PICO , 1, 0) 
datz5$MPB <- rowSums(datz5[,c("MPB.d", "MPB.l")], na.rm=T)
datz5$MPB[is.na(datz5$MPB)==T]<-0

datz5$SB.l <- NA
datz5$SB.d <- NA
datz5$SB.l <- ifelse(datz5$DAMAGE_AGENT_CD1==bb & datz5$SPCD == PIEN | datz5$DAMAGE_AGENT_CD1==11009 & datz5$SPCD == PIEN, 1, 0) 
datz5$SB.d <- ifelse(datz5$AGENTCD==10 & datz5$SPCD == PIEN, 1, 0)  
datz5$SB <- rowSums(datz5[,c("SB.d", "SB.l")], na.rm=T)
datz5$SB[is.na(datz5$SB)==T]<-0
```

#### Calculation of stand characteristics
For each plot, we then calculated
1 - basal area by host species
2 - quadratic mean diameter (QMD)
3 - basal area dominance (% total basal area) by species
4 - presence and severity of bark beetle activity (% total basal area) by bark beetle species

```{r stand characteristics, eval=F}
## BA
datz5$BA <- datz5$DIA*datz5$DIA * pi/(4*144) *datz5$TPA_UNADJ # sq ft per acre
datz5$BA <- datz5$BA * 1/10.7639 * 2.47105/1 # m2 per ha
BA <- aggregate(BA~PLT_CN+SPCD, datz5, FUN=sum, na.rm=T)
BA.dat <- dcast(BA, PLT_CN~SPCD, value.var="BA", drop=F)
BA.dat$BA.total <- rowSums(BA.dat[,-1], na.rm=T)
BA.dat$BA.Allhosts <- rowSums(BA.dat[,as.character(c(PICO, PIEN, ABLA))], na.rm=T)
BA.dat$BA.Otherhosts <- rowSums(BA.dat[,as.character(otherhosts)], na.rm=T)
BA.dat <- BA.dat[, c("PLT_CN", PICO, PIEN, ABLA, "BA.total", "BA.Allhosts", "BA.Otherhosts")]
colnames(BA.dat) <- c("PLT_CN", paste0("BA.", c("PICO", "PIEN", "ABLA", "total", "Allhosts", "Otherhosts")))
BA.dat[is.na(BA.dat)==TRUE] <- 0

## Trees per ha
datz5$TPH <- datz5$TPA_UNADJ * 2.47105 # trees per ha
TPH <- aggregate(TPH~PLT_CN+SPCD, datz5, FUN=sum, na.rm=T)
TPH.dat <- dcast(TPH, PLT_CN~SPCD, value.var="TPH", drop=F)
TPH.dat$TPH.total <- rowSums(TPH.dat[,-1], na.rm=T)
TPH.dat$TPH.Allhosts <- rowSums(TPH.dat[,as.character(c(PICO, PIEN, ABLA))], na.rm=T)
TPH.dat <- TPH.dat[, c("PLT_CN", PICO, PIEN, ABLA, "TPH.total", "TPH.Allhosts")]
colnames(TPH.dat) <- c("PLT_CN", paste0("TPH.", c("PICO", "PIEN", "ABLA", "total", "Allhosts")))
TPH.dat[is.na(TPH.dat)==TRUE] <- 0

## QMD
datz5$DIAcm <- datz5$DIA * 2.54 # convert to cm
qmd <- aggregate(DIAcm~PLT_CN+SPCD, datz5, FUN=function(x){sqrt(mean(x*x,na.rm=T))})
qmd.dat<- dcast(qmd, PLT_CN~SPCD , value.var="DIAcm", drop=F)
qmd.dat<- qmd.dat[,c("PLT_CN", PICO, PIEN, ABLA)]
colnames(qmd.dat) <- c("PLT_CN", paste0("QMD.",c("PICO", "PIEN", "ABLA")))
qmd.dat.hosts <- aggregate(DIAcm~PLT_CN, datz5[datz5$SPCD %in% c(PICO, PIEN, ABLA),], FUN=function(x){sqrt(mean(x*x,na.rm=T))})
colnames(qmd.dat.hosts) <- c("PLT_CN", "QMD.Allhosts")
qmd.dat.all <- aggregate(DIAcm~PLT_CN, datz5, FUN=function(x){sqrt(mean(x*x,na.rm=T))})
colnames(qmd.dat.all) <- c("PLT_CN", "QMD.total")
qmd.dat<-merge(qmd.dat, qmd.dat.hosts, by="PLT_CN", all=T)
qmd.dat<-merge(qmd.dat, qmd.dat.all, by="PLT_CN", all=T)
qmd.dat[is.na(qmd.dat)] <- 0

## SDI (additive)
datz5$SDIi <- (datz5$DIA/10)^1.6
sdi <- aggregate(SDIi~PLT_CN+SPCD, datz5, FUN=sum)
sdi.dat<- dcast(sdi, PLT_CN~SPCD , value.var="SDIi", drop=F)
sdi.dat<- sdi.dat[,c("PLT_CN", PICO, PIEN, ABLA)]
colnames(sdi.dat) <- c("PLT_CN", paste0("SDI.",c("PICO", "PIEN", "ABLA")))
sdi.dat[is.na(sdi.dat)] <- 0

## Trees per plot of bark beetle affected trees
notreesBB.dat <- merge(merge(aggregate(MPB~PLT_CN,datz5, FUN=sum, na.rm=T),  aggregate(SB~PLT_CN,datz5, FUN=sum, na.rm=T), by="PLT_CN", all=T), aggregate(WBBB~PLT_CN,datz5, FUN=sum, na.rm=T), by="PLT_CN", all=T)
colnames(notreesBB.dat) <- c("PLT_CN", paste0("stand.", c("MPB", "SB", "WBBB")))
notreesBB.dat$stand.allBBs <- rowSums(notreesBB.dat[,2:4])
notreesBB.dat[is.na(notreesBB.dat)] <- 0

## Total BA of bark beetle affected trees
BABB.dat <- merge(merge(aggregate(MPB*BA~PLT_CN,datz5, FUN=sum, na.rm=T),  aggregate(SB*BA~PLT_CN,datz5, FUN=sum, na.rm=T), by="PLT_CN", all=T), aggregate(WBBB*BA~PLT_CN,datz5, FUN=sum, na.rm=T), by="PLT_CN", all=T)
colnames(BABB.dat) <- c("PLT_CN", paste0("BA.", c("MPB", "SB", "WBBB")))
BABB.dat$BA.allBBs <- rowSums(BABB.dat[,2:4])
BABB.dat[is.na(BABB.dat)] <- 0

df<- unique(datz5[,c("PLT_CN", "PREV_PLT_CN", "PLOT.STATEAB", "INVYR", "ELEV", "LAT", "LON" ),])
df <-merge(df, TPH.dat, by="PLT_CN", all=T)
df <-merge(df, qmd.dat, by="PLT_CN", all=T)
df <-merge(df, BA.dat, by="PLT_CN", all=T)
df <-merge(df, sdi.dat, by="PLT_CN", all=T)
df <- merge(df, notreesBB.dat, by="PLT_CN", all=T)
df <- merge(df, BABB.dat, by="PLT_CN", all=T)
df[is.na(df$PREV_PLT_CN)==TRUE, "PREV_PLT_CN"] <- -9999
df <- na.omit(df)

df$agentSB <- ifelse(df$stand.SB>0, "SB", "absent")
df$agentMPB <- ifelse(df$stand.MPB>0, "MPB", "absent")
df$agentWBBB <- ifelse(df$stand.WBBB>0, "WBBB", "absent")

df$agents <-  do.call(paste, c(df[c("agentSB", "agentMPB", "agentWBBB")],sep="-"))
df$agents <- factor(df$agents, levels=c("absent-absent-absent","absent-MPB-absent","SB-absent-absent", "SB-MPB-absent", "absent-absent-WBBB", "SB-absent-WBBB", "absent-MPB-WBBB", "SB-MPB-WBBB"), labels=c("none", "MPB", "SB", "MPB & SB", "WBBB", "SB & WBBB", "MPB & WBBB", "MPB, SB & WBBB"))

df$nagents <- df$agents
df$nagents <- factor(df$nagents, levels=c("none", "MPB", "SB", "MPB & SB", "WBBB", "SB & WBBB", "MPB & WBBB", "MPB, SB & WBBB"), labels=c(0, 1, 1, 2, 1, 2, 2, 3))

df$hostsPICO <- ifelse(df$TPH.PICO>0, "PICO", "absent")
df$hostsPIEN <- ifelse(df$TPH.PIEN>0, "PIEN", "absent")
df$hostsABLA <- ifelse(df$TPH.ABLA>0, "ABLA", "absent")
df$hosts <-  do.call(paste, c(df[c("hostsPICO", "hostsPIEN", "hostsABLA")],sep="-"))
df$hosts <- factor(df$hosts, levels=c('PICO-absent-absent','absent-absent-ABLA', 'absent-PIEN-absent', 'PICO-PIEN-absent',   'PICO-absent-ABLA', 'absent-PIEN-ABLA', 'PICO-PIEN-ABLA', 'absent-absent-absent'), labels= c('PICO', 'ABLA', 'PIEN', 'PICO & PIEN', 'PICO & ABLA', 'PIEN & ABLA', 'PICO, PIEN, & ABLA', 'none'), ordered=TRUE)

df$nhosts <- df$hosts
df$nhosts <- factor(df$nhosts, levels=c('PICO', 'ABLA', 'PIEN', 'PICO & PIEN', 'PICO & ABLA', 'PIEN & ABLA', 'PICO, PIEN, & ABLA', 'none'), labels=c(1, 1,1,2,2,2,3,0))

write.csv(df, here("Data/Processed/FIAdat-df.csv"),row.names=F)
```

#### Selection of plots
We selected all plots that were part of the annual inventory and where all subplots were inventoried.

Within this plots, we characterized stand structure and composition using only live and recently killed (i.e. killed within the past ~10 years) trees greater that 12.7 cm DBH within the subplot.

```{r select plots}
df <- read.csv(here("Data/Processed/FIAdat-df.csv"))

##select most recent inventories
df.recent <- df

df.recent <-df.recent[!(df.recent$PREV_PLT_CN %in% df.recent$PLT_CN), ] # select plots that have not been subsequently sampled

##select inventories conducted within the last 10 years
df.recent <- df.recent[df.recent$INVYR %in% 2010:2019,]
df.recent$LONx <- df.recent$LON
df.recent$LATx <- df.recent$LAT
##select inventories with coordinates located in the Intermountain West
df.recent.sf <- st_as_sf(df.recent, coords = c("LONx", "LATx"), crs = 4326)
df.recent.sf <- st_transform(df.recent.sf,  crs = aea.proj)
states <-st_read(here("Data/Spatial/States/cb_2018_us_state_20m.shp"),quiet=T)
states <- st_transform(states, crs = aea.proj)
df.recent.sf.j.states <- st_join(df.recent.sf, states)
df.recent.sf.j.states<- df.recent.sf.j.states[df.recent.sf.j.states$STUSPS %in% c("AZ", "CO", "ID","MT", "NM", "UT", "WY"),]

df.recent.sf <- df.recent.sf.j.states
df.recent <- df.recent.sf
st_geometry(df.recent) <- NULL

write.csv(df.recent, here("Data/Processed/FIAdat-recent.csv"))
```

```{r merge data, eval=F}
df.recent <- read.csv(here("Data/Processed/FIAdat-recent.csv"))
write.csv(df, here("Data/Processed/FIAdat-df-merge.csv"), row.names=F)
```

```{r reorganize by focal host, eval=F}
df <- read.csv(here("Data/Processed/FIAdat-df-merge.csv"))

formatp <- function(d, host="PIEN", agent="SB"){
  df.p <- d[which(d[, paste0("TPH.", host)]>0),]
  #df.p <- d
  df.p$TPH.host <- df.p[, paste0("TPH.", host)]
  df.p$QMD.host <- df.p[, paste0("QMD.", host)]
  df.p$BA.host <- df.p[, paste0("BA.", host)]

  df.p$stand.BB <- df.p[, paste0("stand.", agent)]
  df.p$BA.BB <- df.p[, paste0("BA.", agent)]
  df.p$affected <- ifelse(df.p$stand.BB>0, "affected", "unaffected")
  df.p$focalhost <- host

  df.p$BAD.host <- df.p$BA.host/df.p$BA.total
  df.p$TPHD.host <- df.p$TPH.host/df.p$TPH.total
  #df.p$pressure.BB <- df.p[,paste0("pressure", agent)]

  return(df.p)
}

df <- rbind(rbind(formatp(df), formatp(df, host="PICO", agent="MPB")), formatp(df, host="ABLA", agent="WBBB"), formatp(df, host="Allhosts", agent="allBBs"))

### Proportion of total BA affected
df$propTBA <- df$BA.BB/df$BA.total

### Proportion of host BA affected
df$propHBA <- df$BA.BB/df$BA.Allhosts

### Proportion of total trees affected
df$propTTPH <- (df$stand.BB*unique(6.018046)*2.47105)/df$TPH.total

### Proportion of host trees affected
df$propHTPH <- (df$stand.BB*unique(6.018046)*2.47105)/df$TPH.host

### Affected
df$affected01 <- ifelse(df$stand.BB>0, "affected", "unaffected")

write.csv(df, here("Data/Processed/FIAdat-df-merge-re.csv"),row.names=F)
```

## Analyses

### Descriptive analyses
```{r import processed data}
df <- read.csv( here("Data/Processed/FIAdat-df-merge-re.csv"))
df$nhostsF <-factor(df$nhosts, levels=1:3, labels=c("one", "two", "three"), ordered=T)
df$hosts <- factor(df$hosts, levels=c("PICO", "PIEN", "ABLA", "PICO & PIEN", "PICO & ABLA", "PIEN & ABLA", "PICO, PIEN, & ABLA"), labels=c("PICO", "PIEN", "ABLA", "PICO\n& PIEN", "PICO\n& ABLA", "PIEN \n& ABLA", "PICO, PIEN,\n& ABLA"), ordered=T)
df$hostsFPlong <- factor(df$hosts, levels=c("PICO", "PIEN", "ABLA", "PICO\n& PIEN", "PICO\n& ABLA", "PIEN \n& ABLA", "PICO, PIEN,\n& ABLA"), labels=c("PICO", "PIEN", "ABLA", "PICO & PIEN", "PICO & ABLA", "PIEN & ABLA", "PICO, PIEN, & ABLA"), ordered=T)

df$nagentsF <- factor(df$nagents, levels=0:3, ordered=T)
df$agentsF <- factor(df$agents, levels=c("none", "MPB", "SB", "WBBB", "MPB & SB","MPB & WBBB",  "SB & WBBB", "MPB, SB & WBBB"), labels=c("unaffected", "MPB", "SB", "WBBB", "MPB\n& SB","MPB &\nWBBB",  "SB &\nWBBB", "MPB,\nSB, &\nWBBB"), ordered=T)
df$agentsFplong <- factor(df$agents, levels=c("none", "MPB", "SB", "WBBB", "MPB & SB","MPB & WBBB",  "SB & WBBB", "MPB, SB & WBBB"), labels=c("unaffected", "MPB", "SB", "WBBB", "MPB & SB","MPB & WBBB",  "SB & WBBB", "MPB, SB, & WBBB"), ordered=T)

df$focalhost <- factor(df$focalhost, levels=c("PICO", "PIEN", "ABLA", "Allhosts"), labels=c("PICO", "PIEN", "ABLA", "Any"), ordered=T)

df$focalagent <- factor(df$focalhost, levels= c("PICO", "PIEN", "ABLA", "Any"), labels=c("MPB", "SB", "WBBB", "Any"), ordered=T)

df$propBA.PICO <- df$BA.PICO / df$BA.total *100
df$propBA.PIEN <- df$BA.PIEN / df$BA.total *100
df$propBA.ABLA <- df$BA.ABLA / df$BA.total *100
```

```{r map study area of hosts, eval=F}
f19 <- stars::read_stars(here("Data/Spatial/ITSPM2002-subalpine/f19/w001001.adf"), proxy=F)
f93 <- stars::read_stars(here("/Data/Spatial/ITSPM2002-subalpine/f93/w001001.adf"), proxy=F)
f108 <- stars::read_stars(here("/Data/Spatial/ITSPM2002-subalpine/f108/w001001.adf"), proxy=F)

# reclassify to rasters to presence (1) and absence (0)
f19[f19>0] <- 1 
f93[f93>0] <- 1
f108[f108>0] <- 1

# define areas of multiple hosts
f108f93 <- f108 + f93 # add 
f108f93[f108f93<2]<- 0 # reclassify
f108f93[f108f93==2]<- 1 # reclassify
f108f19 <- f108 + f19
f108f19[f108f19<2]<- 0 # reclassify
f108f19[f108f19==2]<- 1 # reclassify
f93f19 <- f93 + f19
f93f19[f93f19<2]<- 0  # reclassify
f93f19[f93f19==2]<- 1 # reclassify

f108f93f19 <- f108 + f93 + f19

# define areas of one host
f108f93f19re1 <-f108f93f19
f108f93f19re1 [f108f93f19re1 ==1 ]<-1
f108f93f19re1 [f108f93f19re1>1]<-0

# define areas of two hosts
f108f93f19re2 <-f108f93f19
f108f93f19re2 [f108f93f19re2 ==1 ]<-0
f108f93f19re2 [f108f93f19re2==3]<-0 
f108f93f19re2 [f108f93f19re2 ==2 ]<-1

# limit maps of presence (1) and absence (0) to only areas with one host
f108 <- f108 * f108f93f19re1 
f93 <- f93 * f108f93f19re1
f19 <- f19 * f108f93f19re1 
f93[f93==1]<- 2 # give unique labels
f19[f19==1]<- 3 # give unique labels

# create maps of two hosts limited to areas with two hosts
f108f93  <- f108f93 * f108f93f19re2 
f108f93[f108f93==1]<-4 # give unique labels

f108f19 <- f108f19 * f108f93f19re2
f108f19[f108f19==1]<-5 # give unique label

f93f19 <- f93f19 * f108f93f19re2
f93f19[f93f19==1]<-6 # give unique labels

# create maps of all three hosts
f108f93f19[f108f93f19<=2]<-0
f108f93f19[f108f93f19==3]<-7 # give unique labels

fplot <- f108 + f93 + f19 + f108f93 + f108f19 + f93f19+ f108f93f19
st_crs(fplot) <- "EPSG:5070" # assign projection

fplot[fplot==0] <-NA

states <-st_read(here("Data/Spatial/States/cb_2018_us_state_20m.shp"),quiet=T)
IMW <- states[states$STUSPS %in% c("AZ", "CO", "ID","MT", "NM", "UT", "WY"),]
notIMW <- states[!states$STUSPS %in% c("AZ", "CO", "ID","MT", "NM", "UT", "WY"),]
IMW <- st_transform(IMW, "EPSG:5070")
notIMW <- st_transform(notIMW, "EPSG:5070")

ext <- as.matrix(extent(IMW))

sa.tmap <- tm_shape(IMW, bbox=tmaptools::bb(ext)) + tm_fill(col="#f5f5f5")+tm_borders()+tm_shape(fplot) +tm_raster(style = "cat", palette = c("#8c510a", "#bf812d", "#dfc27d", "#80cdc1", "#35978f", "#01665e", "#003c30"), colorNA=NULL, legend.show = TRUE, title="Hosts                    ", labels = c("PICO", "PIEN", "ABLA", "PICO & PIEN", "PICO & ABLA", "PIEN & ABLA", "PICO, PIEN, & ABLA"))+ tm_shape(notIMW) + tm_fill(col="white")+tm_compass(position=c(0.95, 0.88), just=c("center", "bottom"))+tm_layout(legend.bg.color="white")+tm_graticules(lines=FALSE)+tm_layout(legend.outside=TRUE,legend.text.size=0.55)+tm_scale_bar(position = c(0.01, 0.001), just="left", breaks=c(0,100),text.size=1.1)+tm_shape(IMW) +tm_borders()

Fig.file <- here("Results/Figures/Figure-Studyarea.jpg")
jpeg(Fig.file, width=fig.width1.5,height=fig.width1.5,units="in", res=300)
sa.tmap 
whatever <- dev.off()
```

```{r describe frequency of hosts and agents, eval=T}
## Hosts
hosts <- df[!df$focalhost=="Any",] %>% group_by(focalhost) %>% summarise(n = n())%>%mutate(freq = n / sum(n))
hostgroups <- df[df$focalhost=="Any",] %>% group_by(hosts) %>% summarise(n = n())%>%mutate(freq = n / sum(n))

## Affected hosts
affectedhosts <- df[!df$focalhost=="Any",] %>% group_by(focalhost, affected) %>% summarise(n = n())%>% mutate(freq = prop.table(n))
affectednhosts <- df[df$focalhost=="Any",] %>% group_by(nhosts, affected) %>% summarise(n = n())%>% mutate(freq = prop.table(n))
affectednhosts<- affectednhosts[affectednhosts$affected=="affected", ]
affectednhosts<- affectednhosts[order(affectednhosts$freq, decreasing=T),]

affectedhostgroups <- df[df$focalhost=="Any",] %>% group_by(hosts, affected) %>% summarise(n = n())%>% mutate(freq = prop.table(n))
affectedhostgroups<- affectedhostgroups[affectedhostgroups$affected=="affected", ]
affectedhostgroups<- affectedhostgroups[order(affectedhostgroups$freq, decreasing=T),]

## Agents given host presence
agents <- df[!df$focalhost=="Any",] %>% group_by(focalagent, affected) %>% summarise(n = n())%>%
  mutate(freq = n / sum(n))
PICOwMPB <- as.numeric(round(agents[agents$focalagent=="MPB" & agents$affected=="affected", "freq"]*100, 0))
PIENwSB <- as.numeric(round(agents[agents$focalagent=="SB" & agents$affected=="affected", "freq"]*100, 0))
ABLAwWBBB<- as.numeric(round(agents[agents$focalagent=="WBBB" & agents$affected=="affected", "freq"]*100, 0))

## Agents given presence of any agent
agentso <- df[!df$focalhost=="Any" & df$affected=="affected",] %>% group_by(focalagent) %>% summarise(n = n())%>%
  mutate(freq = n / sum(n))

agentsbyhosts <- df[!df$focalhost=="Any",] %>% group_by(focalagent, affected) %>% summarise(n = n())%>% mutate(freq = prop.table(n))

p1 <- ggplot(df[df$focalhost=="Any",], aes(x=hosts, fill=interaction(hosts, affected))) + geom_bar(width=0.75, stat="count", col="black")+ylab("number of plots")+theme(axis.text.x = element_text(angle = 45, hjust=1,vjust=1),legend.position = "none", legend.title=element_blank())+scale_fill_manual(values=c(c("#8c510a", "#bf812d", "#dfc27d", "#80cdc1", "#35978f", "#01665e", "#003c30"), adjust_transparency(c("#8c510a", "#bf812d", "#dfc27d", "#80cdc1", "#35978f", "#01665e", "#003c30"), alpha=0.5)))

Fig.file <- here("Results/Figures/Figure-PresencebyHostcombos.jpg")
jpeg(Fig.file, width=fig.width1,height=fig.width1/1.5,units="in", res=300)
p1
whatever <- dev.off()
```

```{r determine co-occurrence of hosts}
Fig.file <- here("Results/Figures/Figure-AgentbyHostcombo.jpg")
df$hostsFPshort<- df$hostsFPlong
df$hostsFPshort <- factor(df$hostsFPlong, levels=levels(df$hostsFPlong), labels=c("PICO", "PIEN", "ABLA", "PICO\n&\nPIEN", "PICO\n&\nABLA", "PIEN\n&\nABLA", "PICO,\nPIEN,\n&\nABLA"))
df$agentsFplongrev <- df$agentsFplong
df$agentsFplongrev <- factor(df$agentsFplongrev, levels=rev(levels( df$agentsFplong)))
jpeg(Fig.file, width=fig.width2,height=fig.width1.5,units="in",res=300)
ggplot(df[df$focalhost=="Any",], aes(x=hostsFPshort,fill=agentsFplongrev))+geom_bar()+theme(legend.title = element_blank())+xlab("hosts")+scale_fill_manual(values=p2col)
whatever <- dev.off()
```

### Determine suitable hosts
Construct random forest models using synthetic minority oversampling technique (SMOTE)

#### MPB
```{r fit MPB RF model, eval=F}
response <- "agentMPB"
vars <- c("BA.PICO", "QMD.PICO", "propBA.PICO")
datBB.c <- df[df$focalhost=="Any",]

splitit <- initial_split(datBB.c, prop=.7, strata=response)
datz.train <- training(splitit)
datz.test <- testing(splitit)
datz.train.cv <- vfold_cv(datz.train, strata=response, v=10)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train)

# set sample weights
wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set model
rf_model <- rand_forest() %>% set_args( trees=tune(), min_n=tune()) %>% set_engine("ranger", importance = "permutation", probability=T, seed=913, sample.fraction=wts) %>% set_mode("classification")

# set the workflow
rf_workflow <- workflow() %>% add_recipe(BBoccurrence_recipe) %>% add_model(rf_model)

# tune model
tgrid <- expand.grid(trees = c(100,500), min_n=c(10, 50, 100))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy, sens, spec))

tuneres <- rf_tune_results %>% collect_metrics()

param_best<- rf_tune_results %>% select_best(metric = "roc_auc")

# Use backward selection to identify important variables
#dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, ntree=param_best$trees, min.node.size=param_best$min_n)
#registerDoSEQ()
#vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))
vars.final <- vars
BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(trees=param_best$trees, min_n=param_best$min_n, splitrule="gini") %>% set_engine("ranger", importance = "permutation", probability=T, sample.fraction=wts) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit, metrics=metric_set(roc_auc, accuracy, sens, spec)) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set. 
test_performance <- rf_workflow_final_fit %>% collect_metrics()
write.csv(test_performance, here(paste0("Results/RF-testperformance-", response, "-occurrence-2.csv")), row.names=F)

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)

write.csv(confusion$table, here(paste0("Results/RF-confusiontable-", response, "-occurrence-2.csv")), row.names=F)

# variable importance
final_model <- fit(rf_workflow_final, data=datz.train)
ranger_obj <- pull_workflow_fit(final_model)$fit

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "SBpred")
write.csv(x, here(paste0("Results/RF-pred-", response, "-occurrence-2.csv")), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
write.csv(x, here(paste0("Results/RF-vimp-", response, "-occurrence-2.csv")), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class=strsplit(response, split="agent")[[1]][2], prob=T)
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here(paste0("Results/RF-partialdependence-", response, "-occurrence-2.csv")), row.names=F)
```

#### SB
```{r fit SB RF model, eval=F}
response <- "agentSB"
vars <- c("BA.PIEN", "QMD.PIEN", "propBA.PIEN")
datBB.c <- df[df$focalhost=="Any",]

splitit <- initial_split(datBB.c, prop=.7, strata=response)
datz.train <- training(splitit)
datz.test <- testing(splitit)
datz.train.cv <- vfold_cv(datz.train, strata=response, v=10)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train)

# set sample weights
wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set model
rf_model <- rand_forest() %>% set_args( trees=tune(), min_n=tune()) %>% set_engine("ranger", importance = "permutation", probability=T, seed=913, sample.fraction=wts) %>% set_mode("classification")

# set the workflow
rf_workflow <- workflow() %>% add_recipe(BBoccurrence_recipe) %>% add_model(rf_model)

# tune model
tgrid <- expand.grid(trees = c(100,500), min_n=c(10, 50, 100))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy, sens, spec))

tuneres <- rf_tune_results %>% collect_metrics()

param_best<- rf_tune_results %>% select_best(metric = "roc_auc")

# Use backward selection to identify important variables
#dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, ntree=param_best$trees, min.node.size=param_best$min_n)
#registerDoSEQ()
#vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))
vars.final <- vars
BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(trees=param_best$trees, min_n=param_best$min_n, splitrule="gini") %>% set_engine("ranger", importance = "permutation", probability=T, sample.fraction=wts) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit, metrics=metric_set(roc_auc, accuracy, sens, spec)) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set. 
test_performance <- rf_workflow_final_fit %>% collect_metrics()
write.csv(test_performance, here(paste0("Results/RF-testperformance-", response, "-occurrence-2.csv")), row.names=F)

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)

write.csv(confusion$table, here(paste0("Results/RF-confusiontable-", response, "-occurrence-2.csv")), row.names=F)

# variable importance
final_model <- fit(rf_workflow_final, data=datz.train)
ranger_obj <- pull_workflow_fit(final_model)$fit

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "SBpred")
write.csv(x, here(paste0("Results/RF-pred-", response, "-occurrence-2.csv")), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
write.csv(x, here(paste0("Results/RF-vimp-", response, "-occurrence-2.csv")), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class=strsplit(response, split="agent")[[1]][2], prob=T)
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here(paste0("Results/RF-partialdependence-", response, "-occurrence-2.csv")), row.names=F)
```

#### WBBB
```{r fit WBBB RF model, eval=F}
response <- "agentWBBB"
vars <- c("BA.ABLA", "QMD.ABLA", "propBA.ABLA")
datBB.c <- df[df$focalhost=="Any",]

splitit <- initial_split(datBB.c, prop=.7, strata=response)
datz.train <- training(splitit)
datz.test <- testing(splitit)
datz.train.cv <- vfold_cv(datz.train, strata=response, v=10)

# set formula
BBoccurrence_recipe <- recipe(formula(paste(response, "~", paste(vars, collapse = " + "))), datz.train)

# set sample weights
wts <- 1/(table(datBB.c[,response])/ max(table(datBB.c[,response])))
wts <- wts/sum(wts)

# set model
rf_model <- rand_forest() %>% set_args( trees=tune(), min_n=tune()) %>% set_engine("ranger", importance = "permutation", probability=T, seed=913, sample.fraction=wts) %>% set_mode("classification")

# set the workflow
rf_workflow <- workflow() %>% add_recipe(BBoccurrence_recipe) %>% add_model(rf_model)

# tune model
tgrid <- expand.grid(trees = c(100,500), min_n=c(10, 50, 100))
rf_tune_results <- rf_workflow %>% tune_grid(resamples = datz.train.cv,  grid = tgrid, metrics = metric_set(roc_auc, accuracy, sens, spec))

tuneres <- rf_tune_results %>% collect_metrics()

param_best<- rf_tune_results %>% select_best(metric = "roc_auc")

# Use backward selection to identify important variables
#dfo.vsurf <- VSURF::VSURF(datz.train[,vars], factor(datz.train[,response]), RFimplem="ranger", parallel=T, ntree=param_best$trees, min.node.size=param_best$min_n)
#registerDoSEQ()
#vars.final <- unique(c(vars[dfo.vsurf$varselect.pred]))
vars.final <- vars
BBoccurrence_recipe_final <- recipe(formula(paste(response, "~", paste(vars.final, collapse = " + "))), data=datBB.c)

rf_model_final <- rand_forest() %>% set_args(trees=param_best$trees, min_n=param_best$min_n, splitrule="gini") %>% set_engine("ranger", importance = "permutation", probability=T, sample.fraction=wts) %>% set_mode("classification")

rf_workflow_final <- workflow() %>% add_recipe(BBoccurrence_recipe_final) %>% add_model(rf_model_final)
rf_workflow_final_fit<- rf_workflow_final %>% last_fit(splitit, metrics=metric_set(roc_auc, accuracy, sens, spec)) #emulates the process where, after determining the best model, the final fit on the entire training set is needed and is then evaluated on the test set. 
test_performance <- rf_workflow_final_fit %>% collect_metrics()
write.csv(test_performance, here(paste0("Results/RF-testperformance-", response, "-occurrence-2.csv")), row.names=F)

# generate a confusion matrix for test set
test_predictions <- rf_workflow_final_fit %>% collect_predictions()
confusion <- test_predictions %>% conf_mat(response,estimate = .pred_class)

write.csv(confusion$table, here(paste0("Results/RF-confusiontable-", response, "-occurrence-2.csv")), row.names=F)

# variable importance
final_model <- fit(rf_workflow_final, data=datz.train)
ranger_obj <- pull_workflow_fit(final_model)$fit

# generate predictions for entire dataset
x <- datBB.c$PLT_CN
x <- cbind(x, predict(ranger_obj, datBB.c)$predictions[,2])
colnames(x) <- c("PLT_CN", "SBpred")
write.csv(x, here(paste0("Results/RF-pred-", response, "-occurrence-2.csv")), row.names=F)

x <- as.data.frame(ranger_obj$variable.importance)
write.csv(x, here(paste0("Results/RF-vimp-", response, "-occurrence-2.csv")), row.names=F)

# partial dependence
pd <- NULL
for(z in vars.final){
  pdz <- partial(ranger_obj, train=datz.train, pred.var=z, which.class=strsplit(response, split="agent")[[1]][2], prob=T)
  pdz$var <- z
  colnames(pdz) <- c("value","yhat","var")
  pd <- rbind(pd, pdz)
}
write.csv(pd,  here(paste0("Results/RF-partialdependence-", response, "-occurrence-2.csv")), row.names=F)
```

```{r determine suitable stands}
MPBpred <- read.csv(here("Results/RF-pred-agentMPB-occurrence.csv"))
SBpred <- read.csv(here("Results/RF-pred-agentSB-occurrence.csv"))
WBBBpred <- read.csv(here("Results/RF-pred-agentWBBB-occurrence.csv"))

BBpred <- merge(merge(MPBpred, SBpred, by="PLT_CN"), WBBBpred, by="PLT_CN")
colnames(BBpred) <- c("PLT_CN", "MPBpred", "SBpred", "WBBBpred")

threshold <-0.5
BBpred$MPBsuitable <- ifelse(BBpred$MPBpred > threshold, "MPB", "absent")
BBpred$SBsuitable <- ifelse(BBpred$SBpred > threshold, "SB", "absent")
BBpred$WBBBsuitable <- ifelse(BBpred$WBBBpred >threshold, "WBBB", "absent")

BBpred$agentsuitable <-  do.call(paste, c(BBpred[c("SBsuitable", "MPBsuitable", "WBBBsuitable")],sep="-"))
BBpred$agentsuitable <- factor(BBpred$agentsuitable, levels=c("absent-absent-absent","absent-MPB-absent","SB-absent-absent", "SB-MPB-absent", "absent-absent-WBBB", "SB-absent-WBBB", "absent-MPB-WBBB", "SB-MPB-WBBB"), labels=c("none", "MPB", "SB", "MPB & SB", "WBBB", "SB & WBBB", "MPB & WBBB", "MPB, SB & WBBB"))

BBpred$suitablehosts <- factor(BBpred$agentsuitable, levels=c("none", "MPB", "SB", "WBBB", "MPB & SB",  "MPB & WBBB", "SB & WBBB",  "MPB, SB & WBBB"), labels=c("none", "PICO", "PIEN", "ABLA","PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ordered=T)

```

## Host co-occurrence

### Given suitable stand conditions, is co-occurrence of suitable hosts common?
```{r determine co-occurrence of suitable stands}
bbo.dat<- merge(BBpred, df, id.vars="PLT_CN", all=T)
bbo.dat <- bbo.dat[bbo.dat$focalagent=="Any",]
cocount <-  as.data.frame(table(bbo.dat$agentsFplongrev))
MPBco <-sum(cocount[cocount$Var1 %in% c("MPB & SB",  "MPB & WBBB","MPB, SB & WBBB"),]$Freq)
SBco <-sum(cocount[cocount$Var1 %in% c("MPB & SB",  "SB & WBBB","MPB, SB & WBBB"),]$Freq)
WBBBco <-sum(cocount[cocount$Var1 %in% c("MPB & WBBB",  "SB & WBBB","MPB, SB & WBBB"),]$Freq)

cocount <-  as.data.frame(table(bbo.dat$suitablehosts))
PICOco <-sum(cocount[cocount$Var1 %in% c("PICO & PIEN",  "PICO & ABLA","PICO, PIEN & ABLA"),]$Freq)
PIENco <-sum(cocount[cocount$Var1 %in% c("PICO & PIEN",  "PIEN & ABLA","PICO, PIEN & ABLA"),]$Freq)
ABLAco <-sum(cocount[cocount$Var1 %in% c("PICO & ABLA",  "PIEN & ABLA","PICO, PIEN & ABLA"),]$Freq)

codat <- data.frame(agent =c( "MPB", "SB", "WBBB"), agentco=c(MPBco, SBco, WBBBco), hostco=c(PICOco, PIENco, ABLAco))

bbo.dat$hostsFPshort<- bbo.dat$hostsFPlong
bbo.dat$hostsFPshort <- factor(bbo.dat$hostsFPlong, levels=levels(bbo.dat$hostsFPlong), labels=c("PICO", "PIEN", "ABLA", "PICO\n&\nPIEN", "PICO\n&\nABLA", "PIEN\n&\nABLA", "PICO,\nPIEN,\n&\nABLA"))
bbo.dat$agentsFplongrev <- bbo.dat$agentsFplong
bbo.dat$agentsFplongrev <- factor(bbo.dat$agentsFplongrev, levels=rev(levels( bbo.dat$agentsFplong)))

ggfreq <- ggplot(bbo.dat[bbo.dat$focalhost=="Any",], aes(x=hostsFPshort,fill=agentsFplongrev))+geom_bar()+theme(legend.title = element_blank())+xlab("hosts")+scale_fill_manual(values=p2col)

ggsave( here("Results/Figures/FigureAgentfrequencybyhosts.jpg"), plot=ggfreq,   width=fig.width2,height=fig.width1.5,units="in")

```

### Is occurrence more likely or severity greater in stands with multiple hosts?

```{r occurrence by suitable hosts}
dfpred <- merge(BBpred, df, id.vars="PLT_CN", all=T)
mpb <- as.data.frame(table(dfpred[dfpred$focalagent=="MPB","affected"], dfpred[dfpred$focalagent=="MPB","suitablehosts"]))
sb <- as.data.frame(table(dfpred[dfpred$focalagent=="SB","affected"], dfpred[dfpred$focalagent=="SB","suitablehosts"]))
wbbb <- as.data.frame(table(dfpred[dfpred$focalagent=="WBBB","affected"], dfpred[dfpred$focalagent=="WBBB","suitablehosts"]))
any<- as.data.frame(table(dfpred[dfpred$focalagent=="Any","affected"], dfpred[dfpred$focalagent=="Any","suitablehosts"]))

results <- NULL
for(j in c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA")){
  anyj <- any[any$Var2==j,]
  mpbj <- mpb[mpb$Var2==j,]
  sbj <- sb[sb$Var2==j,]
  wbbbj <- wbbb[wbbb$Var2==j,]

  x <- prop.test(x=c(mpbj[1,3],anyj[1,3]), n=c(mpbj[1,3]+mpbj[2,3], anyj[1,3]+anyj[2,3]))

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "MPB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)

  x <- prop.test(x=c(sbj[1,3],anyj[1,3]), n=c(sbj[1,3]+sbj[2,3], anyj[1,3]+anyj[2,3]))

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "SB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)

   x <- prop.test(x=c(wbbbj[1,3],anyj[1,3]), n=c(wbbbj[1,3]+wbbbj[2,3], anyj[1,3]+anyj[2,3]), alternative="less")

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA, prop1=NA, prop2 =NA, lower=NA, upper=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "WBBB"
  res$X <- x$statistic
  res$p <- x$p.value
  res$prop1 <- x$estimate[1]
  res$prop2 <- x$estimate[2]
  res$lower <- x$conf.in[1]
  res$upper <- x$conf.in[2]
  x <-NULL
  results <- rbind(results, res)
}

results$p <- round(results$p,3)

x <- dfpred[is.na(dfpred$suitablehosts)==FALSE,]
x <- x[x$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"),]
x[x$suitablehosts == "PICO & PIEN" & x$focalagent=="WBBB","affected"]<- 0
x[x$suitablehosts == "PICO & ABLA" & x$focalagent=="SB","affected"]<- 0
x[x$suitablehosts == "PIEN & ABLA" & x$focalagent=="MPB","affected"]<- 0

xx <- x %>% group_by(suitablehosts,focalagent) %>% count(affected)
xxx <- x %>% group_by(suitablehosts) %>% count()

xxxx <- merge(xx, xxx, by="suitablehosts")
xxxx$prop <- xxxx$n.x/xxxx$n.y
xxxx <- xxxx[xxxx$affected=="affected",]
xxxx <- xxxx[,c(1,2,6)]
colnames(xxxx) <- c("suitablehosts", "barkbeetle", "prop")

resultsx <- merge(results, xxxx, id.vars=c("suitablehosts", "barkbeetle"), all=T)
resultsx$barkbeetle <- factor(resultsx$barkbeetle, levels=c("MPB", 'SB', "WBBB", "Any"), ordered=T)
resultsx <- resultsx[is.na(resultsx$prop)==FALSE,]
resultsx$suitablehosts <- factor(resultsx$suitablehosts, levels=c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), labels=c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ordered=T)
resultsx$pplot <- ifelse(resultsx$p<0.05, "*", "")
resultsx$pplot[is.na(resultsx$pplot) ==TRUE] <- ""

ggoccurrence <- ggplot(resultsx, aes(x=barkbeetle,y=prop, fill=barkbeetle))+geom_bar(stat="identity")+theme(legend.title = element_blank())+geom_text(aes(label=pplot), vjust=-0.25,size=6)+facet_grid(.~suitablehosts, scales="free_x", space="free_x")+xlab("bark beetle species")+scale_fill_manual(values=c(p1col[1:3], gray(0.3)))+theme(legend.position = "none")+ylab("proportion of plots")

ggsave( here("Results/Figures/FigureOccurrencebySuitablehosts.jpg"), plot=ggoccurrence,  width=fig.width2, height=fig.width1,units="in")

```

```{r severity by suitable hosts}
dfpred <- merge(BBpred, df, id.vars="PLT_CN", all=T)
mpb <- dfpred[dfpred$focalagent=="MPB" & dfpred$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA",   "PICO, PIEN, & ABLA"), ]
sb <- as.data.frame(table(dfpred[dfpred$focalagent=="SB","affected"], dfpred[dfpred$focalagent=="SB","suitablehosts"]))
wbbb <- as.data.frame(table(dfpred[dfpred$focalagent=="WBBB","affected"], dfpred[dfpred$focalagent=="WBBB","suitablehosts"]))
any<- as.data.frame(table(dfpred[dfpred$focalagent=="Any","affected"], dfpred[dfpred$focalagent=="Any","suitablehosts"]))

dfpred <- dfpred[dfpred$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ]

dfpred <- dfpred[dfpred$affected01=="affected",]

results <- NULL
for(j in c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA")){

  x <- kruskal.test(propTBA~focalagent, data=dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent %in% c('Any', 'MPB'),])

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "MPB"
  res$X <- x$statistic
  res$p <- x$p.value
  x <-NULL
  results <- rbind(results, res)

    x <- kruskal.test(propTBA~focalagent, data=dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent %in% c('Any', 'SB'),])

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "SB"
  res$X <- x$statistic
  res$p <- x$p.value
  x <-NULL
  results <- rbind(results, res)

      x <- kruskal.test(propTBA~focalagent, data=dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent %in% c('Any', 'WBBB'),])

  res <- data.frame(suitablehosts=NA, barkbeetle=NA, X=NA, p=NA)
  res$suitablehosts <-j
  res$barkbeetle <- "WBBB"
  res$X <- x$statistic
  res$p <- x$p.value
  x <-NULL
  results <- rbind(results, res)

}

results$pplot <- ifelse(results$p<0.05, "*", NA)

dfpred$pplot <- NA
for(j in c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA")){

  x <- dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'MPB',]
  xplt <- x[x$propTBA==max(x$propTBA), "PLT_CN"][1]
  dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'MPB' &dfpred$PLT_CN==xplt, ]$pplot <- results[results$suitablehosts==j & results$barkbeetle=='MPB', "pplot"]

    x <- dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'SB',]
  xplt <- x[x$propTBA==max(x$propTBA), "PLT_CN"][1]
  dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'SB' &dfpred$PLT_CN==xplt, ]$pplot <- results[results$suitablehosts==j & results$barkbeetle=='SB', "pplot"]

    x <- dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'WBBB',]
  xplt <- x[x$propTBA==max(x$propTBA), "PLT_CN"][1]
  dfpred[dfpred$suitablehosts ==j  & dfpred$focalagent == 'WBBB' & dfpred$PLT_CN==xplt, ]$pplot <- results[results$suitablehosts==j & results$barkbeetle=='WBBB', "pplot"]

}

dfpred <- dfpred[!(dfpred$suitablehosts == "PICO & PIEN" & dfpred$focalagent=="WBBB"),]
dfpred <- dfpred[!(dfpred$suitablehosts == "PICO & ABLA" & dfpred$focalagent=="SB"),]
dfpred <- dfpred[!(dfpred$suitablehosts == "PIEN & ABLA" & dfpred$focalagent=="MPB"),]

ggseverity <- ggplot(dfpred, aes(x=focalagent, y=propTBA*100, fill=focalagent, colour=focalagent))+geom_boxplot(notch=T)+facet_grid(.~suitablehosts, scales="free_x", space="free_x")+xlab("bark beetle species")+scale_fill_manual(values=adjust_transparency(c(p1col[1:3], gray(0.3)), 0.5))+scale_colour_manual(values=c(p1col[1:3], gray(0.3)))+theme(legend.position = "none")+ylim(1,100)+geom_text(aes(label=pplot), vjust=-0.25,size=10, col="black")+ylab("Severity (% BA affected)")

ggsave(here("Results/Figures/FigureSevertitybySuitablehosts.jpg"), plot=ggseverity, width=fig.width2,height=fig.width1,units="in")

```

## Agent co-occurrence

### Given suitable stand conditions, is co-occurrence of multiple agents common?

### What are the effects of multiple agents?
```{r severity by suitable hosts}
dfpred <- merge(BBpred, df, id.vars="PLT_CN", all=T)
dfpred <- dfpred[dfpred$suitablehosts %in% c("PICO & PIEN",  "PICO & ABLA", "PIEN & ABLA",  "PICO, PIEN, & ABLA"), ]

dfpred <- dfpred[dfpred$affected01=="affected",]

dfpred <- dfpred[!(dfpred$suitablehosts=="PICO & PIEN" & dfpred$agentsF %in% c("WBBB", "MPB &\nWBBB", "SB &\nWBBB", "MPB,\nSB, &\nWBBB")),]

dfpred <- dfpred[!(dfpred$suitablehosts=="PICO & ABLA" & dfpred$agentsF %in% c("SB", "MPB\n& SB", "SB &\nWBBB", "MPB,\nSB, &\nWBBB")),]

dfpred <- dfpred[!(dfpred$suitablehosts=="PIEN & ABLA" & dfpred$agentsF %in% c("MPB", "MPB\n& SB", "MPB &\nWBBB", "MPB,\nSB, &\nWBBB")),]

ggmultipelseverity <- ggplot(dfpred[dfpred$focalagent=="Any",], aes(x=agentsF, y=propTBA*100, fill=agentsF))+geom_boxplot()+facet_grid(.~suitablehosts, scales="free_x", space = "free_x")+theme(legend.position="none")

ggsave(here("Results/Figures/FigureSeverityMultiple.jpg"), plot=ggmultipelseverity , width=fig.width2,height=fig.width1,units="in")
```

# Results

### Patterns

Across the Intermountain West, the most commonly occurring species was lodgepole pine, which was present in `r as.numeric(round(hosts[hosts$focalhost=="PICO", "freq"]*100,0))`% of plots, followed by subalpine fir (`r as.numeric(round(hosts[hosts$focalhost=="ABLA", "freq"]*100,0))`% of plots), and Engelmann spruce (`r as.numeric(round(hosts[hosts$focalhost=="PIEN", "freq"]*100,0))`% of plots). Given the presence of the focal host, the most commonly occurring bark beetle species was the MPB, which was found in `r PICOwMPB`% of plots with lodgepole pine. SB and WBBB, were found at much lower rates; SB was found in `r PIENwSB`% of plots with Engelmann spruce and WBBB was found in `r ABLAwWBBB`% of plots with subalpine fir.

The most commonly occurring tree community included only lodgepole pine (n=`r as.numeric(hostgroups[hostgroups$hosts=="PICO", "n"])`), followed by plots with all three hosts (n=`r as.numeric(hostgroups[hostgroups$hosts=="PICO, PIEN,\n& ABLA", "n"])`) and Engelmann spruce-sublapine fir plots (n=`r as.numeric(hostgroups[hostgroups$hosts=="PIEN \n& ABLA", "n"])`) (Figs. \@ref(fig:FigSA) -  \@ref(fig:FigPresHosts)). Logepole pine-Engelmann spruce communities were the least common (n=`r as.numeric(hostgroups[hostgroups$hosts=="PICO\n& PIEN", "n"])`), followed by plots with only subalpine fir (n=`r as.numeric(hostgroups[hostgroups$hosts=="ABLA", "n"])`) and only Engelmann spruce ((n=`r as.numeric(hostgroups[hostgroups$hosts=="PIEN", "n"])`)).

Consistent with diversity begets diversity hypothesis, the relative occurrence of bark beetle activity was greatest in stands with all three hosts (Fig \@ref(fig:FigPresHosts); `r as.numeric(round(affectednhosts[affectednhosts$nhosts==3, "freq"] *100, 0))`% of plots with all three hosts were affected by bark beetles). Stands with only one host were least likely to experience bark beetle infestation and stands with (`r as.numeric(round(affectednhosts[affectednhosts$nhosts==1, "freq"] *100, 0))`% of plots with only one host) and stands with two hosts were intermediate  (`r as.numeric(round(affectednhosts[affectednhosts$nhosts==2, "freq"] *100, 0))`% of plots with two hosts).

However host richness also varied with other factors known to influence outbreak presence and severity, notably stand structure and composition (Fig. \@ref(fig:FigStandSC)), climate (Fig. \@ref(fig:FigClimate)), and landscape scale patterns of host availability and bark beetle pressure (Fig. \@ref(fig:FigPressure)).

```{r FigSA, fig.cap = "The distribution of host species presence across the study area. Data are from the Individual Tree Species Atlas (Ellenwood et al. 2015) and represent conditions in ca. 2002.", out.width = "322.56pt"}
knitr::include_graphics( here("Results/Figures/Figure-Studyarea.jpg"))
```

## Co-occurrence of agents

```{r FigCoOccurrence, fig.cap = "Co-occurrence",  out.width = "504.8pt"}

knitr::include_graphics( here("Results/Figures/FigureAgentfrequencybyhosts.jpg"))
```

Few stands were affected by multiple agents.

### Is occurrence more likely or severity greater in stands with multiple hosts?

```{r FigOccurrence, fig.cap = "Severity",  out.width = "504.8pt"}

knitr::include_graphics( here("Results/Figures/FigureOccurrencebySuitablehosts.jpg"))
```

```{r FigSeverity, fig.cap = "Severity", out.width = "504.8pt"}

knitr::include_graphics( here("Results/Figures/FigureSevertitybySuitablehosts.jpg"))
```

# Discussion

# References

<div id="refs"></div>

## Supplement
### Randon Forest Modeling
```{r FigPartial,  fig.cap = "A caption", out.width = "504.8pt"}
MPBpp <- read.csv(here("Results/RF-partialdependence-agentMPB-occurrence.csv"))
MPBpp$focalagent <- "MPB"
SBpp <- read.csv(here("Results/RF-partialdependence-agentSB-occurrence.csv"))
SBpp$focalagent <- "SB"
WBBBpp <- read.csv(here("Results/RF-partialdependence-agentWBBB-occurrence.csv"))
WBBBpp$focalagent <- "WBBB"

ppdat <- rbind(rbind(MPBpp, SBpp), WBBBpp)
ppdat$var[ppdat$var %in% paste0("BA.", c("PICO", "PIEN", "ABLA"))] <- "BA"
ppdat$var[ppdat$var %in% paste0("QMD.", c("PICO", "PIEN", "ABLA"))] <- "QMD"
ppdat$var[ppdat$var %in% paste0("propBA.", c("PICO", "PIEN", "ABLA"))] <- "propBA"

x <- df[!df$focalagent=="Any",]
x$propBA <- x$BA.host/x$BA.total
x <- x[,c("QMD.host", "propBA", "BA.host", "focalagent", "affected01")]
x$affected01 <- ifelse(x$affected01=="affected", 1, 0)

xm <- melt(x, id.vars=c("focalagent", "affected01"))
colnames(xm) <- c("focalagent", "affected01", "var", "value")
xm$var <- factor(xm$var, levels=c("QMD.host", "propBA", "BA.host"), labels=c("QMD", "propBA", "BA"))

pBA1 <- ggplot(ppdat[ppdat$var=="BA",], aes(x=value, y=yhat, col=focalagent))+geom_smooth()+scale_color_manual(values=p1col)+theme(legend.position="none")+ylab("probability of occurrence")+theme(axis.title.x=element_blank(),axis.text.x=element_blank() )

pBA2 <- ggplot(xm[xm$var=="BA",],aes(x=value, y=focalagent, fill=interaction(focalagent, affected01)))+geom_boxplot()+scale_fill_manual(values=c(adjust_transparency(p1col[1:3], alpha=0.25), p1col[1:3]))+theme(legend.position="none")+ylab("bark beetle")+xlab("basal area")

pQMD1 <- ggplot(ppdat[ppdat$var=="QMD",], aes(x=value, y=yhat, col=focalagent))+geom_smooth()+scale_color_manual(values=p1col)+theme(legend.position="none")+ylab("probability of occurrence")+theme(axis.title.x=element_blank(),axis.text.x=element_blank() )

pQMD2 <- ggplot(xm[xm$var=="QMD",],aes(x=value, y=focalagent, fill=interaction(focalagent, affected01)))+geom_boxplot()+scale_fill_manual(values=c(adjust_transparency(p1col[1:3], alpha=0.25), p1col[1:3]))+theme(legend.position="none")+ylab("")+xlab("QMD")

ppropBA1 <- ggplot(ppdat[ppdat$var=="propBA",], aes(x=value, y=yhat, col=focalagent))+geom_smooth()+scale_color_manual(values=p1col)+theme(legend.position="none")+ylab("probability of occurrence")+theme(axis.title.x=element_blank(),axis.text.x=element_blank() )

ppropBA2 <- ggplot(xm[xm$var=="propBA",],aes(x=value, y=focalagent, fill=interaction(focalagent, affected01)))+geom_boxplot()+scale_fill_manual(values=c(adjust_transparency(p1col[1:3], alpha=0.25), p1col[1:3]))+theme(legend.position="none")+ylab("")+xlab("propBA")

ppplot <- pBA1  + pQMD1 +theme(axis.title.y=element_blank(),axis.text.y=element_blank()) +ppropBA1 +theme(axis.title.y=element_blank(),axis.text.y=element_blank())+ pBA2 + pQMD2 +theme(axis.text.y=element_blank())  + ppropBA2 +theme(axis.text.y=element_blank()) + plot_layout(ncol=3, heights=c(2,1))

ggsave(here("Results/Figures/PartialDependence.jpg"), plot=ppplot, width=fig.width2,height=fig.width1,units="in")
knitr::include_graphics(here("Results/Figures/PartialDependence.jpg"))

```

### Comparison of stand structure and composition, climate, and regional bark beetle activity  
```{r FigStandSC,  fig.cap = "A caption", out.width = "322.56pt", eval=F}
datBB.c <-df.sub[!df.sub$focalhost=="Any",]

QMD.melt <- melt(datBB.c[,c("hosts",paste0("QMD.", c("PIEN", "PICO", "ABLA")))], "hosts")
QMD.melt$variable <- factor(QMD.melt$variable, levels=paste0("QMD.", c("PICO","PIEN",  "ABLA")), labels=c("PICO", "PIEN", "ABLA"), ordered=T)
QMD.melt$facet <- "QMD"

TPH.melt <- melt(datBB.c[,c("hosts",paste0("TPH.", c("PIEN", "PICO", "ABLA")))], "hosts")
TPH.melt$variable <- factor(TPH.melt$variable, levels=paste0("TPH.", c("PICO","PIEN",  "ABLA")), labels=c("PICO", "PIEN", "ABLA"), ordered=T)
TPH.melt$facet <- "TPH"

BA.melt <- melt(datBB.c[,c("hosts",paste0("BA.", c("PIEN", "PICO", "ABLA")))], "hosts")
BA.melt$variable <- factor(BA.melt$variable, levels=paste0("BA.", c("PICO","PIEN",  "ABLA")), labels=c("PICO", "PIEN", "ABLA"), ordered=T)
BA.melt$facet <- "BA"

propBA.melt <-datBB.c[,c("hosts",paste0("BA.", c("PIEN", "PICO", "ABLA", "total")))]
propBA.melt$propBA.PIEN <- propBA.melt$BA.PIEN / propBA.melt$BA.total *100
propBA.melt$propBA.ABLA <- propBA.melt$BA.ABLA / propBA.melt$BA.total *100
propBA.melt$propBA.PICO <- propBA.melt$BA.PICO / propBA.melt$BA.total *100

propBA.melt <-  melt(propBA.melt[,c("hosts",paste0("propBA.", c("PIEN", "PICO", "ABLA")))], "hosts")
propBA.melt$variable <- factor(propBA.melt$variable, levels=paste0("propBA.", c("PICO","PIEN",  "ABLA")), labels=c("PICO", "PIEN", "ABLA"), ordered=T)
propBA.melt$facet <- "BA dominance (%)"

plotdat <- rbind(propBA.melt, TPH.melt, BA.melt, QMD.melt)

ggss <- ggplot(plotdat[plotdat$value>0,], aes(x=hosts, y=value,fill=variable))+geom_boxplot(outlier.size = 0.2,color="black",lwd=0.2,position = position_dodge(preserve = "single"))+scale_fill_brewer(palette ="BuPu")+ylab("value")+xlab("")+facet_wrap(.~facet, scales="free_y", ncol=1)+theme(legend.position = "bottom", legend.title=element_blank())

ggsave(here("Results/Figures/Figure-HostStructureByhostIdentity.jpg"), plot=ggss , width=fig.width1.5, height=fig.width1*2, units="in")
knitr::include_graphics(here("Results/Figures/Figure-HostStructureByhostIdentity.jpg"))
```

```{r generate maps, eval=F}
bbo.dat<- merge(BBpred, df, id.vars="PLT_CN", all=T)

df.all <- bbo.dat[bbo.dat$focalhost == "Any", ]

df.all.sf <- st_as_sf(df.all, coords = c("LON", "LAT"), crs = 4326)

states <-st_read(here("Data/Spatial/States/cb_2018_us_state_20m.shp"),quiet=T)
IMW <- states[states$STUSPS %in% c("AZ", "CO", "ID","MT", "NM", "UT", "WY"),]

ext <- as.matrix(extent(IMW))
#ext[1,1] <- ext[1,1] -0.25
#ext[1,2] <- ext[1,2] +0.25
#ext[2,1] <- ext[2,1] -1
#ext[2,2] <- ext[2,2] +1

map1 <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="PICO",]) +tm_dots(col="agentsFplong", legend.show = FALSE,size=.0025, palette=rev(p2col)) +tm_compass(position=c(0.9, 0.02), just=c("center", "bottom"), text.size=0.5, size=2)+tm_layout(main.title = "PICO",main.title.size=0.5,main.title.position="center")

map2 <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="PIEN",]) +tm_dots(col="agentsFplong", legend.show = FALSE,size=.0025, palette=rev(p2col)) +tm_layout(main.title = "PIEN",main.title.size=0.5,main.title.position="center")

map3 <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="ABLA",]) +tm_dots(col="agentsFplong", legend.show = FALSE,size=.0025, palette=rev(p2col)) +tm_layout(main.title = "ABLA",main.title.size=0.5, main.title.position="center")

map4 <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="PICO & PIEN",]) +tm_dots(col="agentsFplong", legend.show = FALSE,size=.0025, palette=rev(p2col)) +tm_layout(main.title = "PICO & PIEN",main.title.size=0.5,main.title.position="center")

map5 <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="PICO & ABLA",]) +tm_dots(col="agentsFplong", legend.show = FALSE,size=.0025, palette=rev(p2col)) +tm_layout(main.title = "PICO & ABLA", main.title.size=0.5, main.title.position="center")

map6 <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="PIEN & ABLA",]) +tm_dots(col="agentsFplong", legend.show = FALSE,size=.0025, palette=rev(p2col)) +tm_layout(main.title = "PIEN & ABLA", main.title.size=0.5, main.title.position="center")

map7 <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="PICO, PIEN, & ABLA",]) +tm_dots(col="agentsFplong", legend.show = FALSE,size=.0025, palette=rev(p2col))+tm_layout(main.title = "PICO, PIEN, & ABLA", main.title.size=0.5, main.title.position="center")

legend.map <- tm_shape(IMW, bbox=tmaptools::bb(ext)) +  tm_fill(col=grey(level=0.95)) +  tm_borders()  + tm_shape(df.all.sf[df.all.sf$hostsFPlong=="PICO, PIEN, & ABLA",]) +tm_dots(col="agentsFplong", size=4, title="bark beetle species",  palette=rev(p2col))+tm_layout(legend.only = TRUE,legend.text.size = 0.5,legend.title.size = 0.65)+tm_scale_bar(position = c(0.5,0.16), text.size=0.5, just="center", breaks=c(0,250,500))

Fig.file <- here("Results/Figures/Figure-Map.jpg")
jpeg(Fig.file, width=fig.width1.5, height=fig.width2/2, units="in")
tmap_arrange(map1, map2,map3, map4, map5, map6, map7,legend.map, nrow=2)
whatever <- dev.off()

```

```{r FigMap, fig.cap="a caption", out.width = "504.8pt"}
knitr::include_graphics(here("Results/Figures/Figure-Map.jpg"))
```